---
format: 
  elsevier-html: default
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super

editor: visual
title: |
  DNA metabarcoding diet analysis in reindeer is quantitative 
  and integrates feeding over several weeks

# if short-title is defined, then it's used
short-title: Ouups
bibliography: bibliography/bibliography.bib
copyright: 
  holder: Reinin project
  year: 2024
author:
  - name: Stefaniya Kamenova
    email: stefaniya.kamenova@gmail.com
    attributes:
      corresponding: true
    orcid: 0000-0002-7241-0506
    affiliations:
      - ref: cees
      - ref: denrm
      - ref: nmhmsofia
  - name: Pernille Meyer
    affiliations:
      - ref: cees
  - name: Anne Krag Brysting
    affiliations:
      - ref: cees
  - name: Leo Rescia
    affiliations:
      - ref: damb
  - name: Lars P. Folkow
    affiliations:
      - ref: damb
  - name: Galina Gusarova
    email: galina.gusarova@uit.no
    role: "Equal contribution"
    affiliations:
      - ref: cees
      - ref: damb
      - ref: aumn
      - ref: dbsp
  - name: Eric Coissac
    email: eric.coissac@metabarcoding.org
    role: "Equal contribution"
    orcid: 0000-0001-7507-6729
    affiliations:
      - ref: leca

affiliations:
  - id: cees
    name: |
      Centre for Ecological and Evolutionary Synthesis, 
      University of Oslo
    city: Oslo
    country: Norway
  - id: denrm
    name: |
      Departments of Ecology and Natural Resource Management, 
      Norwegian University of Life Sciences
    city: Ås
    state: Norway
  - id: nmhmsofia
    name: |
      National Museum of Natural History,
      Bulgarian Academy of Sciences
    city: Sofia
    state: Bulgaria
  - id: damb
    name: |
      Department of Arctic and Marine Biology, 
      The Arctic University of Norway
    city: Tromsø
    state: Norway
  - id: aumn
    name: |
      The Arctic University Museum of Norway, 
      The Arctic University of Norway
    city: Tromsø
    state: Norway
  - id: dbsp
    name: |
      Department of Botany, 
      St Petersburg State University
    city: St Petersburg
    state: Russia
  - id: leca
    name: |
      Laboratoire d'Ecologie Alpine, 
      CNRS, Université Grenoble Alpes
    city: Grenoble
    state: France
    
abstract: |
  Filtering of the EUKA02 DNA metabarcoding raw data.
---

# Setting up the R environment

## Install missing packages

```{r}
packages <- c("igraph", "tidyverse", "devtools","R.utils","vegan")
install.packages(setdiff(packages,
                         rownames(installed.packages())),
                 dependencies = TRUE
                )
```


## Loading of the R libraries

- `ROBITools` package is used to read result files produced by OBITools.

- `ROBITaxonomy` package provides function allowing to query OBITools formated taxonomy.

```{r loads_robitools, warning=FALSE, message=FALSE}
if (!"ROBITools" %in% rownames(installed.packages())) {
  # ROBITools are not available on CRAN and have to be installed
  # from http://git.metabarcoding.org using devtools

  metabarcoding_git <- "https://git.metabarcoding.org/obitools"

  devtools::install_git(paste(metabarcoding_git,
                              "ROBIUtils.git",
                              sep="/"))

  devtools::install_git(paste(metabarcoding_git,
                              "ROBITaxonomy.git",
                              sep="/"))
  devtools::install_git(paste(metabarcoding_git,
                              "ROBITools.git",
                              sep="/"))
}

library(ROBITools)
library(ROBITaxonomy)
```


- `tidyverse` [@Wickham2019-ra] provides various method for efficient data manipulation and plotting via `ggplot2` [@Wickham2016-vl]

```{r loads_tidyverse, warning=FALSE, message=FALSE}
library(tidyverse)
```

```{r Rutils, warning=FALSE, message=FALSE}
library(R.utils)
```

```{r Vegan, warning=FALSE, message=FALSE}
library(vegan)
```

```{r magrittr, warning=FALSE, message=FALSE}
library(magrittr)
```


```{r}
source("methods.R")
```

# Loading the data

## Load the NCBI taxonomy

```{r}
if (! file.exists("Data/ncbi20210212.adx")) {
  gunzip("Data/ncbi20210212.adx.gz",remove=FALSE)
  gunzip("Data/ncbi20210212.ndx.gz",remove=FALSE)
  gunzip("Data/ncbi20210212.rdx.gz",remove=FALSE)
  gunzip("Data/ncbi20210212.tdx.gz",remove=FALSE)
}
```

```{r}
taxo <- read.taxonomy("Data/ncbi20210212")
```



## Loading the metabarcoding data

```{r}
if (! file.exists("Data/Rawdata/EUKA02_all_paired.ali.assigned.ann.diag.uniq.ann.c1.l10.clean.EMBL.tag.ann.sort.uniq.grep.tab"))
    gunzip("Data/Rawdata/EUKA02_all_paired.ali.assigned.ann.diag.uniq.ann.c1.l10.clean.EMBL.tag.ann.sort.uniq.grep.tab.gz",remove=FALSE)


EUKA02.raw = import.metabarcoding.data("Data/Rawdata/EUKA02_all_paired.ali.assigned.ann.diag.uniq.ann.c1.l10.clean.EMBL.tag.ann.sort.uniq.grep.tab")
```

## Loading the metadata

```{r}
samples.metadata = read_csv("Data/Faeces/sampling_dates.csv", 
                            show_col_types = FALSE)
```

# Sample description

Normalization of samples names

```{r include=FALSE}
sample_names = sub("PCR_NC", "PCRNC", samples(EUKA02.raw)$sample)
sample_names = sub("PCR_POS", "PCRPOS", sample_names)
sample_names = sub("DNANC1", "DNANC_1", sample_names)

samples(EUKA02.raw)$sample
```

Extract information relative to PCR replicates and sample names.

```{r}
sample_names_split = strsplit(as.character(sample_names), "_R")

replicate = sapply(sample_names_split, function(x) x[length(x)])
sample_id = sapply(sample_names_split, function(x) x[1])

samples_desc = data.frame(name = samples(EUKA02.raw)$sample, replicate = replicate, sample_id = sample_id)

EUKA02.raw@samples = samples_desc
EUKA02.raw@motus <- EUKA02.raw@motus %>% select(-starts_with("obiclean_status:"))
```


## Categorize MOTUs

DNA Sequence of the synthetic sequence used as EUKA02 positive controls.

```{r}
Standard1 = "taagtctcgcactagttgtgacctaacgaatagagaattctataagacgtgttgtcccat"
```

 - Identify which MOTU is corresponding to the positive control sequence 
   and associated it to category `standard1`.
 - All the MOTUs exhibiting a similarity with one of the reference SPER01
   database greater than 80% is tagged as `EUKA02`
 - The remaining sequences are tagged as `Unknown`


```{r}
sequence_type = rep("Unknown", nrow(motus(EUKA02.raw)))
sequence_type[which(motus(EUKA02.raw)$`best_identity:db_EUKA`> 0.80)] = "EUKA02"
sequence_type[which(motus(EUKA02.raw)$sequence == Standard1)] = "standard1"

EUKA02.raw@motus$sequence_type = as.factor(sequence_type)
table(EUKA02.raw@motus$sequence_type)
```

```{r}
spermatophyta.taxid <- ecofind(taxo,patterns = "^Spermatophyta$")
lecanoromycetidae.taxid = ecofind(taxo,"^Lecanoromycetidae$")

to_keep = (is.subcladeof(taxo,EUKA02.raw@motus$taxid,spermatophyta.taxid) | 
             EUKA02.raw@motus$taxid == spermatophyta.taxid) |
           (is.subcladeof(taxo,EUKA02.raw@motus$taxid,lecanoromycetidae.taxid) | 
              EUKA02.raw@motus$taxid == lecanoromycetidae.taxid)

table(to_keep)
```

```{r}
EUKA02.plant_lichen <- EUKA02.raw[,which(to_keep)]
```


# Curation procedure

## Select motus occuring at least at 1% in at least one PCR

```{r}
norare = apply(decostand(reads(EUKA02.plant_lichen),method = "total"),
               MARGIN = 2,
               FUN = max) >= 0.01
table(norare)
```

```{r}
EUKA02.norare <- EUKA02.plant_lichen[,which(norare)]

```


## Filtering for PCR outliers

Only library 1 and 2 have individually tagged PCR replicates

```{r}
library_3.ids = read.csv("Data/samples_library_3.txt",
                         stringsAsFactors = FALSE,
                         header = FALSE)[,1]
```

```{r}
library3.keep = gsub("_R.?$","_R",rownames(EUKA02.norare)) %in% library_3.ids
EUKA02.lib3 = EUKA02.norare[library3.keep,]
EUKA02.lib12= EUKA02.norare[!library3.keep,]

dim(EUKA02.lib3)
dim(EUKA02.lib12)

```

Load the script containing the selection procedure implemented in function 
`tag_bad_pcr`.

```{r}
source("Select_PCR.R")
```

### First selection round

```{r}
keep1 = tag_bad_pcr(samples = samples(EUKA02.lib12)$sample_id, 
                   counts = reads(EUKA02.lib12),
                   plot = TRUE,
                   threshold=0.2
                  )
```
Histogram shows the empirical distribution of the PCR replicate distances. The red vertical dashed line indicates the threshold used to discard outlier PCRs. The green vertical dashed line indicates the mode of the observed distribution.

```{r}
table(keep1$keep)
```

`FALSE`is the count of PCR to discard, `TRUE` the count of PCR conserved at the end of this selection round.

```{r}
samples(EUKA02.lib12)$name[!keep1$keep]
```
Above is the list of the ids of the discarded PCRs.


```{r}
EUKA02.lib12.k1 = EUKA02.lib12[keep1$keep,]
```

### Second selection round

```{r}
keep2 = tag_bad_pcr(samples = samples(EUKA02.lib12.k1)$sample_id, 
                   counts = reads(EUKA02.lib12.k1),
                   plot = TRUE,
                   threshold=0.2
                  )

```

```{r}
table(keep2$keep)
```

```{r}
samples(EUKA02.lib12.k1)$name[!keep2$keep]
```


```{r}
EUKA02.lib12.k2 = EUKA02.lib12.k1[keep2$keep,]
```

### Third selection round

```{r}
keep3 = tag_bad_pcr(samples = samples(EUKA02.lib12.k2)$sample_id, 
                   counts = reads(EUKA02.lib12.k2),
                   plot = TRUE,
                   threshold=0.2
                  )

```

```{r}
table(keep3$keep)
keep3[!keep3$keep,]
```


```{r}
EUKA02.lib12.k3 = EUKA02.lib12.k2[keep3$keep,]
```


### Merge remaining PCR replicates

```{r}
freq = decostand(reads(EUKA02.lib12.k3),
                 method = "total")
EUKA02.lib12.k3$count = reads(EUKA02.lib12.k3)
EUKA02.lib12.k3@reads = freq

EUKA02.merged = aggregate(EUKA02.lib12.k3, MARGIN = 1, by = list(sample_id=samples(EUKA02.lib12.k3)$sample_id), FUN = mean)
```

## Merge lib 1,2 and 3



### Remove controls in library 3

Look for controls left in library 1 and 2

```{r}
rownames(EUKA02.merged)
```


## Remove controls in library 3

```{r}
rownames(EUKA02.lib3)
```

```{r}
EUKA02.lib3.samples = EUKA02.lib3[-(1:10),]
rownames(EUKA02.lib3.samples@reads) = sub("_R.?$","",rownames(EUKA02.lib3.samples))
rownames(EUKA02.lib3.samples)
```

### Merge library 1, 2 and 3

```{r}

EUKA02.lib123.reads = rbind(EUKA02.merged@reads,
                            decostand(EUKA02.lib3.samples@reads,method = "total"))

common = intersect(names(EUKA02.merged@samples),
                   names(EUKA02.lib3.samples@samples))

EUKA02.lib123.samples = rbind(EUKA02.merged@samples[,common],
                          EUKA02.lib3.samples@samples[,common])

EUKA02.lib123 = metabarcoding.data(reads = decostand(EUKA02.lib123.reads,method = "total"),
                               samples = EUKA02.lib123.samples,
                               motus = EUKA02.merged@motus)

dim(EUKA02.lib123)
```



```{r}
EUKA02.lib123@samples$animal_id = sapply(EUKA02.lib123@samples$sample_id, 
                                                function(x) strsplit(as.character(x),"_")[[1]][1])
```

### Check for empty MOTUs


```{r}
zero = colSums(reads(EUKA02.lib123)) == 0
table(zero)
```

```{r}
EUKA02.nozero = EUKA02.lib123[,!zero]
```


# Filter out rare species

```{r}
plot(EUKA02.nozero@motus$`best_identity:db_EUKA`,
     apply(reads(EUKA02.nozero),2,max),
     col=as.factor(EUKA02.nozero@motus$sequence_type),
     log="y",
     ylab="max read fequency in a sample",
     xlab="best identity with the reference database")
abline(h=0.01,col="red",lty=2)
abline(v=0.95,col="red",lty=2)
```
  

```{r}
EUKA02.merged3 = EUKA02.lib123[, apply(reads(EUKA02.lib123),2,max) > 0.01]
```


```{r}
plot(EUKA02.merged3$motus$`best_identity:db_EUKA`,
     apply(reads(EUKA02.merged3),2,max),
     col=as.factor(EUKA02.merged3$motus$sequence_type),
     log="y",
     ylab="max read fequency in a sample",
     xlab="best identity with the reference database")
abline(h=0.01,col="red",lty=2)
abline(v=0.95,col="red",lty=2)
```

## Keep only MOTUs Strictly identical to one of the reference sequence

### First level stringency filter (95% identity)

```{r}
EUKA02.merged4 = EUKA02.merged3[,EUKA02.merged3@motus$`best_identity:db_EUKA` > 0.95]
EUKA02.merged4@reads = decostand(EUKA02.merged4@reads,method = "total")
EUKA02.merged4@motus <- EUKA02.merged4@motus %>% select(-starts_with("obiclean_status:"))
```

```{r}
plot(EUKA02.merged4$motus$`best_identity:db_EUKA`,
     apply(reads(EUKA02.merged4),2,max),
     col=as.factor(EUKA02.merged4$motus$sequence_type),
     log="y",
     ylab="max read fequency in a sample",
     xlab="best identity with the reference database")
abline(h=0.01,col="red",lty=2)
```

### High stringency filtering

```{r}
spermatophyta.taxid <- ecofind(taxo,patterns = "^Spermatophyta$")
EUKA02.merged4@motus$is_spermatophyta <- is.subcladeof(taxo,EUKA02.merged4@motus$taxid,spermatophyta.taxid)
table(EUKA02.merged4@motus$is_spermatophyta)
EUKA02.merged4@motus %>% filter(!is_spermatophyta)
```


```{r}
to_keep <- EUKA02.merged4@motus$`best_identity:db_EUKA` > 0.95
table(to_keep)
EUKA02.merged4@motus %>% filter(!to_keep)
```

```{r}
EUKA02.final <- EUKA02.merged4[,which(to_keep)]
EUKA02.final@reads <- decostand(EUKA02.final@reads,method = "total")
```


# Saving the filtered dataset


## Updating the sample metadata

### Adding samples metadata

```{r}
metadata <- read_csv("Data/Faeces/metadata.csv",
                     show_col_types = FALSE)
```


```{r}
EUKA02.final@samples %<>%
  select(sample_id,animal_id) %>%
  left_join(metadata,by = "sample_id") %>%
  mutate(id = sample_id) %>%
  column_to_rownames("id") %>%
  select(sample_id,animal_id,Sample_number,Date,Sample_time,times_from_birch, Fed_biomass)
```

### Homogenize time from burch

Adds :
  - 6 hours to animal X,
  - 3 hours to animal Y,
  - 4 hours to animal 2

```{r}
EUKA02.final@samples %<>% 
  mutate(times_from_birch = times_from_birch + 
           ifelse(animal_id == "X",6,
         ifelse(animal_id == "Y",3,4)))
```

```{r}
EUKA02.final@samples %<>%
  mutate(Animal_id = ifelse(animal_id == "X","9/10",
         ifelse(animal_id == "Y","10/10","12/10")))
```

## Add MOTUs Metadata 

```{r}
EUKA02.final@motus %<>%
  mutate(category = ifelse(is.subcladeof(taxo,taxid,spermatophyta.taxid),
                           "Plant",
                           "Lichen"))
```

## Only keep samples

```{r}
EUKA02.final <- EUKA02.final[which(str_detect(EUKA02.final@samples$sample_id,"^[XYZ]")),]
```

## Updating count statistics

```{r}
EUKA02.final %<>% 
  update_motus_count() %>%
  update_samples_count() %>%
  clean_empty()
```


## Write CSV files


```{r}
write_csv(EUKA02.final@samples,
          file = "Data/Faeces/FE.Eukaryota.samples.samples.csv")
write_csv(EUKA02.final@motus,
          file = "Data/Faeces/FE.Eukaryota.samples.motus.csv")
write_csv(EUKA02.final@reads %>%
            decostand(method = "total") %>%
            as.data.frame()%>%
            rownames_to_column("id"),
          file = "Data/Faeces/FE.Eukaryota.samples.reads.csv")
```




















